\documentclass[10pt, a4paper]{article}
\clubpenalty1000000
\widowpenalty1000000
\usepackage{lrec2014}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}
\usepackage{listings}
\usepackage{amsmath,amssymb,colortbl,xcolor}
\usepackage{tipa}

\usepackage{fontspec}
\newfontfamily\free{FreeSans}
\usepackage{zhspacing}
\zhspacing
\setmainfont{FreeSans}

\title{Concepticon: A Resource for the Linking of Concept Lists}

\name{Johann-Mattis List$^{1}$, Michael Cysouw$^{2}$, Robert Forkel$^{3}$}

\address{ 
               $^{1}$CRLAO/EHESS and AIRE/UPMC, Paris,
	       $^{2}$Forschungszentrum Deutscher Sprachatlas, Marburg,\\
	       $^{3}$Max Planck Institute for the Science of Human History, Jena
	       \\
	       \href{mailto:mattis.list@lingpy.org}{mattis.list@lingpy.org}, 
	       \href{mailto:email}{emailpending}, 
	       \href{mailto:email}{emailpending}\\
	       }


\abstract{
We present an attempt to link the large amount of different concept lists (aka ``Swadesh lists'')
which are used in the linguistic literature. This resource, the \emph{Concepticon}
(\url{http://concepticon.clld.org}), links
\conceptnumber concepts labels from \listnumber conceptlists to \setnumber concept sets. Each
concept set is given a unique numerical identifier, a unique label, and a human-readable definition.
Concept sets are further structured by defining different relations between the concepts.  The
resource can be used for various purposes. Serving as a rich reference for new
and existing databases in diachronic and synchronic linguistics, it allows researchers a quick
access to studies on semantic change, cross-linguistic polysemies, and semantic associations. 
\newline \Keywords{concepts, concept lists, Swadesh lists, cross-linguistically linked data}}


% this command serves to get easy access to the basis in hyperref
\newcommand{\concepticon}[1]{http://concepticon.clld.org/contributions/#1}
\newcommand{\conceptnumber}{\textcolor{red}{xxx} }
\newcommand{\listnumber}{\textcolor{red}{xxx} }
\newcommand{\setnumber}{\textcolor{red}{xxx} }
\newcommand{\relationnumber}{\textcolor{red}{xxx} }

\begin{document}

\maketitleabstract

\section{Introduction}
In 1950, Morris Swadesh (1909 -- 1967) proposed the idea that certain parts of the lexicon of human
languages are universal, stable over time, and rather resistant to borrowing. As a result, he
claimed that these
parts of the lexicon, which was later called \emph{basic vocabulary}, would be very
useful to address the problem of subgrouping in historical linguistics:
\begin{quote}
[...] it is a well known fact that certain types of morphemes are relatively stable. Pronouns and
numerals, for example, are occasionally replaced either by other forms from the same language or by
borrowed elements, but such replacement is rare. The same is more or less true of other everyday
expressions connected with concepts and experiences common to all human groups or to the groups
living in a given part of the world during a given epoch. \cite[157]{Swadesh1950}
\end{quote}
He illustrated this by proposing a first \emph{list of basic concepts}, which was, in fact, nothing
else than a collection of concept labels, as shown below:\footnote{This list contains 123 items in
total. According to Swadesh, these items occurred both in his original test list of English items,
and in the data on the Salishan languages, which he employed for his first glottochronological
study.}
\begin{quote}
I, thou, he, we, ye, one, two, three, four, five,
six, seven, eight, nine, ten, hundred, all,
animal, ashes, back, bad, bark, belly, big,
black, blood, bone, brother (elder), child
(son or daughter), cloud, cold, come, cry
(weep), dance, day, dog, dust, ear, earth, eat,
egg, eye, far, father, fire, flower, fog, foot,
good, grass, green, hair, hand, head, heart,
here, hit (with fist), hunt, husband, ice,
lake, laugh, leaf, left hand, leg, liver, long,
louse, man, meat, mother, mountain, mouth,
name, near, neck, night, nose, person, rain,
red, right hand, road (trail), root, rope, salt,
sand, short, sing, sister (elder), skin, sky,
small, smoke, snake, snow, speak, spear
(war), star, stone, sun, swim, tail, that, there,
this, tongue, tooth, tree, warm, water, what,
where, white, who, wife, wind, woman, year,
yellow. \cite[161]{Swadesh1950}
\end{quote}
In the following years, Swadesh refined his original concept lists of basic vocabulary items,
thereby reducing the original test list of 215 items first to 200 \cite{Swadesh1952} and then to 100
items \cite{Swadesh1955}. Scholars working on different language families and different datasets
provided further modifications, be it that the concepts which Swadesh had proposed were lacking
proper translational equivalents in the languages they were working on, or that they
turned out to be not as stable and universal as Swadesh had claimed \cite{Matisoff1978,Alpher1999}. 
Up to today, dozens of different concept list have been compiled for various purposes.
They are are used as heuristical tools for the detection of deep genetic
relationships among languages \cite{Dolgopolsky1964}, 
as basic values for traditional
lexicostatistical and glottochronological studies \cite{Dyen1992,Starostin1991}, 
or as litmus test
for dubious cases of language relationship which might be due to inheritance or borrowing 
\cite{McMahon2005b,Chen1996,Wang2006}. 
 
Apart from concept lists proposed for the application in historical linguistics, there is
furthermore a large amount of not explicitly diachronic data, including concept lists serving as the
basis for field work in specific linguistic areas \cite{Kraft1981}, concept lists which serve as the
basis for large surveys on specific linguistic phenomena \cite{Wold2009}, or concept lists which deal with
the internal \emph{structuring} of concepts, be it cognitive associations \cite{Nelson2004},
cross-linguistic polysemies \cite{List2014f}, or frequently recurring semantic shifts
\cite{Bulakh2013}. 
\section{Concept Lists}
Concept lists are -- simply speaking -- lists of concepts. In these lists, concepts are ideally
described with help of a \emph{concept label} and also a short \emph{definition}. Most published
concept lists, however, only contain a concept label. On the other hand, certain concept lists have
been further expanded by
adding structure, such as \emph{rankings}, \emph{divisions}, or \emph{relations}.

\subsection{Purpose of Concept Lists}
Concept lists are compiled for a variety of different
purposes. A major distinction can be made between those concept lists which have been compiled for
the
purpose of \emph{language comparison} and those which have been compiled for the purpose of \emph{concept
comparison}. 
Among the former, we can further distinguish those lists which are used to prove \emph{language
relationship} \cite{Dolgopolsky1964}, those which are used for \emph{linguistic subgrouping}
\cite{Norman2003,Starostin1991,Swadesh1955}, and those which can be used to identify \emph{contact
layers} \cite{Chen1996}. Among the latter, we can distinguish between concept lists with a primarily
\emph{synchronic objective} \cite{Hill2014}, and those with a primarily \emph{diachronic objective}
\cite{Wold2009,Bulakh2013}.

\subsection{Structure of Concept Lists}
The purpose for which a given concept list was originally defined has an intermediate influence on
its structure. Given the multitude of use cases in both synchronic and diachronic linguistics, it is
difficult to give an exhaustive and unique classification schema for all concept lists which have
been compiled in the past. In Table \ref{tab:conceptlists}, we have nevertheless tried to
distinguish eight basic types of concept lists and give one list for each of the types as a
prototypical example.\footnote{For further information regarding these concept lists, just click on
the links in the ``Example" field of the table.}

\begin{table}[h]
\resizebox{0.475\textwidth}{!}{%
\tabular{|p{4cm}|p{4cm}|p{4cm}|}
\hline
\bf Type                                & \bf Example                                  & \bf Purpose \\\hline\hline
basic vocabulary list (``Swadesh list") & \href{\concepticon/Swadesh-1952-200}{Swadesh 1952 / 200 items}                     & subgrouping \\\hline
subdivided concept list                 & \href{\concepticon/Yakhontov-1991-100}{Yakhontov 1991 (see
Starostin 1991) / 35 + 65 items}               & genetic relationship, layer identification \\\hline
``ultra-stable" concept list            & \href{\concepticon/Dolgopolsky-1964-15}{Dolgopolsky 1964 /
15 items}                & genetic relationship \\\hline
questionnaire                           & \href{\concepticon/Allen-2007-500}{Allen 2007 / 500
items}                    & dialect / language comparison \\\hline
ranked list                             & \href{\concepticon/Starostin-2007-110}{Starostin 2007 /
110 items}                  & subgrouping, layer identification \\\hline
list of concept relations               & \href{\concepticon/Bulakh-2013-870}{DatSemShift, Bulakh et
al. 2013 / 2424 items} & representation of concept
relations \\\hline
special-purpose concept list            & \href{\concepticon/Matisoff-1978-200}{Matisoff 1978 / 200
items}                 & subgrouping of Tibeto-Burman languages \\\hline
historical concept list                 & \href{\concepticon/Leibniz-1768-128}{Leibniz 1768 / 128
items}                    & language comparison \\\hline
\endtabular}
\caption{Examples for different types of concept list as they can be found in the literature.}
\label{tab:conceptlists}
\end{table}
\nocite{Swadesh1952,Starostin1991,Dolgopolsky1964,Allen2007,Bulakh2013,Matisoff1978,Leibniz1768}

\section{Linking Concept Lists}
\textcolor{red}{Here, we should describe the basic characteristics of the concepticon, like the way
the things are linked with each other. Maybe, including a graphic would also be useful}
The concepticon is an attempt to link the many different concept lists (``Swadesh Lists'') which are
used in the linguistic literature. In practice, all entries from the various concept lists are
linked to a \emph{concept set} as an intermediate way to reference the concepts. The Concepticon
currently links \conceptnumber concepts from \listnumber concept lists to \setnumber concept sets
and defines \relationnumber relations between the concept sets.
 
A concept list is a collection of concepts that is deemed interesting by scholars. Minimally, it
consists of an \emph{identifier} for each concept which the lists contains, and a \emph{label} by
which the concept is referenced. The creator of a concept list is called a \emph{compiler}. Each
concept list is tight to one or more \emph{sources}, it is given in one or more \emph{source
languages} and was compiled for one or more \emph{target languages}. A \emph{description} gives
further information on each concept list in free, exclusively human-readable form.
 
To facilitate our workflow and to guarantee the comparability of concept lists even if they do not
share concepts which are directly linked via our concept sets, we define additional and very simple
\emph{concept relations} between concept sets (\emph{broader}, \emph{narrower}, \emph{similar}).
Even if the concepts in two or more concept lists are not assigned to the same concept set, they can
still be assigned to concept sets via concept relations.

\section{Examples}
\textcolor{red}{Examples may be useful to be included, they could, however, also be put into a nice
graphic in which the web-application is presented along with the underlying graphs showing the
relations between the concepts}
\subsection{Rain}

\textcolor{red}{Maybe give rain as an example here, as in the slides}

\subsection{Child}

\textcolor{red}{Child example, for hierarchies}

\subsection{Burn}
\textcolor{red}{Burn example to show problems with transitivity etc.}

\section{Using the Concepticon}
\textcolor{red}{HIer eventuell zeigen, wie das Concepticon bei Dictionaria und Lexibank benutzt
werden kann}

\section{Outlook}
\textcolor{red}{heir noch mal sagen, dass wir natürlich noch weiter daran arbeiten.}


\section{Acknowledgements}
\textcolor{red}{vielleicht nicht nötig im Moment...}

\free
\bibliographystyle{lrec2014}
\bibliography{concepticon}

\end{document}

